---
title: Simple Linear Regression"
author: "Arwa Elnasri"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
 
# # Upload libraries
```{r}
library(readr)
library(ggplot2)
library(tidyverse)
library(caTools)


```

# # Upload data
```{r}
setwd("C:/Users/arwae/OneDrive/Documents/Machine learning/Supervised_learning/linear Reg/Arwa'sgithub-linearREG")

dataset <- read_csv("Salary_Data.xls")


```
# # Clean data
```{r}
# check if there is any NA
  any(is.na(dataset))
# There is no NA in the dataset

```


#Explore the data
```{r}
summary(dataset)

hist(dataset$YearsExperience)
hist(dataset$Salary)
# No Outliers or Influential Points

correlation<- cor(dataset$YearsExperience,dataset$Salary)
# There is a high correlation between the two variables

# Visualizing Dataset
dataset %>%
  ggplot(aes(YearsExperience, Salary))+
  geom_point()

# There is a linear relationship between the independent and dependent variables.

```


# Split data
```{r}
set.seed(123)
splited <- sample.split(dataset$Salary, SplitRatio = 0.8)
train_data<-subset(dataset, splited=="TRUE")
test_data<- subset(dataset, splited=="FALSE")  


```

# Apply logistic regression and Analyze the model (MODEL1)
```{r}
Linear_model <- lm(train_data$Salary ~., train_data)

#Analyse the linear model
summary(Linear_model)

hist(linear_m$residuals)



```
# # interpretations of coefficients:
The residuals are approximately normally distributed

Assessing the coefficients of the model:
The coefficients are: intercept= 26995.6, slope =  9152.3
There is a positive relationship between X and Y
pr(>|t|)=the probability that slope is 9152.3 where it should be zero= p-value) < 2e-16 
Since pr<0.05, the predictor variable is significant in predicting the response variable.

# # Assessing the Accuracy of the Model:
R-squared: The predictor can explain 96.47% of the variation size
p_value: it suggests that the overall model is significant.


#Visualize training dataset with linear model
``` {r}
train_data %>%
  ggplot(aes(YearsExperience, Salary))+
  geom_point()+
  geom_line(aes(YearsExperience,y=predict(Linear_model,train_data)), col="red", size=1)+
  labs( title="YearsExperience VS Salary Training data", 
        subtitle= " Linear Model for Training data")

```
# applying the model to the test data
``` {r}
predicted_test<- predict(Linear_model, test_data)

#Visualize the test data with linear model
test_data %>% 
  ggplot(aes(YearsExperience, Salary))+
  geom_point()+
  geom_line(aes(x=YearsExperience, y=predicted_test),col="blue",size=1)+
  labs(title= "YearsExperience VS predicted_test in Train Data",
        subtitle="Linear model in test data")
```

# Analyze the accuracy of the Linear MODEL
```{r}

# make a prediction for each X
predictedY<- predict(Linear_model, test_data)
summary(predictedY)
# Add predictedY to test_data to compare with actual
test_data$predictedSalary <- predictedY
# display the predictions
test_data %>%
  ggplot(aes(x=YearsExperience, y= predictedY))+
  geom_point()

#Evaluate the model by calculating the Root Mean Square Error (RMSE) between the predicted and actual to evaluate the model.

# Create a function to calculate Root Mean Square Error (RMSE)
rmse <- function(error)
  
{
  sqrt(mean(error^2))
}


# Compute error between predicted and actual
error <- (test_data$Salary - test_data$predictedSalary)  

predictionRMSE <- rmse(error)
sqrt(mean((test_data$Salary - test_data$predictedSalary)^2))

mae <- mean(abs(test_data$Salary - test_data$predictedSalary))
mse <- mean((test_data$Salary - test_data$predictedSalary)^2)
rmse <- sqrt(mse)

# Print evaluation metrics
cat("Mean Absolute Error (MAE):", mae, "\n")
cat("Mean Squared Error (MSE):", mse, "\n")
cat("Root Mean Squared Error (RMSE):", rmse, "\n")

```




# # Results Analysis:
The linear pattern in these plots suggests a linear relationship between "YearsExperience" and "Salary"
From visualizing the test data with the linear model, the test data points are relatively closed to the linear model.

Assessing the coefficients of the model:
The coefficients are: intercept= 26995.6, slope =  9152.3
There is a positive relationship between "YearsExperience" and "Salary"
pr(>|t|)=the probability that slope is 9152.3 where it should be zero= p-value) < 2e-16 
Since pr<0.05, the predictor variable is significant in predicting the response variable.

# # Assessing the Accuracy of the Model:
R-squared: The predictor can explain 96.47% of the variation size
p_value: it suggests that the overall model is significant.