---
title: "Assignmentsvm"
author: "Group"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load libraries
```{r}
library(readxl)
library(e1071)
library(tidyverse)
library(ggplot2)
library(caret)
```

# Upload data
```{r}
setwd("C:/Users/arwae/OneDrive/Documents/Machine learning/Supervised_learning/SVM/svm_data")
zoo <- read.csv("zoo.csv")
head(zoo)
str(zoo)
class_data<- read.csv("class.csv")
head(class_data)
str(class_data)

```

# add class to zoo data
```{r}

colnames(class_data) <- c("class_type", 
                          "no_of_animal_species_in_class", 
                          "class_name", 
                          "animal_name"
)

class_data <- class_data %>%
  select(class_type, class_name)


zoo_data <- zoo %>% 
  left_join( class_data, by = "class_type") 
 
  
head(zoo_data)

```


# Data cleaning
```{r}
any(is.na(zoo_data))
# There is no missing values
```

# Data conversion
```{r}
str(zoo_data)
summary(zoo_data)

# Convert into categorical data except class_name
zoo_data[,-1]<- lapply(zoo_data[ ,-1], factor)




# Count the number of unique values for each variable in the log_data dataframe
sapply(zoo_data, function(x) length(unique(x)))


```

#plot the outcome variable
```{r}
zoo_data%>% 
  ggplot(aes(class_name, fill=class_name ))+
  geom_bar()+
  labs(title="Class Names VS Count in Zoo Data",
       x=" Class Name",
       y= "Count")+
  theme_minimal()



```


# Build classification model
## Divide zoo_data into Train and test 

```{r}
str(zoo_data)
# Select integer variables
zoo_data_selected <- zoo_data %>%
  select(-animal_name, -class_name)

set.seed(123)
data_partition <- createDataPartition(y=zoo_data_selected$class_type, p=0.7, list=F)
zoo_train <- zoo_data_selected[data_partition,]
zoo_test <- zoo_data_selected[-data_partition,]

cat("----- Training batch ------")
table(zoo_train$class_type)

cat("----- Testing batch -----")
table(zoo_test$class_type)


```



# Tune to select c and gamma
```{r}

# Here I tried to apply the same function in Shaila DASH WORKSHOP but it doesn't work
# maybe someone could figure out why? and fix it
set.seed(3409)
tune.obj <- tune.svm(class_type~., data = zoo_train, gamma = 2^(-1:1), cost = 2^(2:4))

summary(tune.obj)
plot(tune.obj)

#select c and gamma that give us the least error

```
Select cost= 4 and gamma =0.5 as these values generate the lowest error
we will need a non-linear kernel to specify the model. We will be using the radial kernel 
# Build SVM model
```{r}

# Using radial kernel support vector classifier.

svm_model <- svm(class_type ~ ., 
                 data=zoo_train,
                 kernel = "radial",
                 cost= 4,
                 gamma= 0.5,
                 probability = TRUE)




summary(svm_model)


```

# Predict class type in test data
```{r}

# predicting class type in test data based on the training model

zoo_test$predicted <- predict(svm_model, zoo_test)
head(zoo_test)
```

# Chech the accuracy of the model
```{r}
# Confusion matrix
con_matrix <- confusionMatrix(zoo_test$class_type, zoo_test$predicted)
print(con_matrix)

```
 The model has 49 support vectors.

In the confusion matrix, The model was able to classify all the animals correctly except one in class 3 which represent Reptile , resulting in an accuracy of 96.43%. 






