---
title: "Logistic Reg GitHub"
author: "Arwa Elnasri"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Upload libraries
```{r, include = FALSE}
library(ggplot2)
library(caret)
library(tidyverse)
library(corrplot)

```

## Upload data
```{r}


log_data<-read.csv("train.csv")
test_data<-read.csv("test.csv")

```
## Clean data
```{r}
#check if any column has any missing values
apply(log_data, 2, function(x) any(is.na(x)))
apply(test_data, 2, function(x) any(is.na(x)))


# Age has missing values, will fill it with mean of age, Fare in the test_data also has missing values.
log_data$Age[is.na(log_data$Age)]<- mean(log_data$Age, na.rm = T)
test_data$Age[is.na(test_data$Age)]<- mean(test_data$Age, na.rm = T)
test_data$Fare[is.na(test_data$Fare)]<- mean(test_data$Fare, na.rm = T)
```

## Checking data type and data conversion 
```{r}
#str(log_data)
#str(test_data)

#log_data[,c(2, 3, 5, 12)] <- lapply(log_data[,c(2, 3, 5, 12)], factor)
#test_data[, c(2, 4 , 11)] <- lapply(test_data[,c(2, 4 , 11)], factor)

log_data[,c(2, 3, 5,7, 12)] <- lapply(log_data[,c(2, 3, 5,7, 12)], factor)
test_data[, c(2, 4,6, 11)] <- lapply(test_data[,c(2, 4,6 , 11)], factor)

str(log_data)
str(test_data)
```
## Explore the data
```{r}
log_data %>% 
  ggplot(aes(Survived, fill=Sex))+
  geom_bar()+
  theme_minimal()+
  labs(title="survived vs Sex")


set.seed(123)
log_data %>%
  ggplot(aes(Age,Survived, col=Sex))+
  geom_jitter(height = 0.05)+
  labs(title = " Age vs survived in Titanic")

  
```


## Select variables
```{r}

#explore cat vairables
#(prop.table(table(log_data$Survived, log_data$Sex),2)*100)"

## Select variables
# Select variable that could affect the classification output

log_data_selected <- log_data[, c(2,3,5,6,7,8,10,12)]

```
## Split data
```{r}
set.seed(123)
spliting_index <- createDataPartition(log_data_selected$Survived, p=0.75, list = FALSE)
train_data <- log_data_selected[spliting_index, ]
validating_data <- log_data_selected[- spliting_index, ]

## check that the proportion of spam is similar
prop.table(table(train_data$Survived))
prop.table(table(validating_data$Survived ))
```

## Apply logistic regression and analyze the model
```{r}
log_model1 <- glm(train_data$Survived ~., train_data, family = binomial)

summary(log_model1)
```
## Interpretations of coefficients:
AIC= 601.76
1. Sexmale is negative (reference : female) indicating that being male decrease the probability of survival compare to female.
2.Pclass has 3 levels. Pclass3 is negative (reference : Pclass1) indicating that being in Pclass3 decrease the probability of survival compare to Pclass1.
3. Age is negative indicating that being older decrease the probability of survival.
4.Embarked, Parch and Fare have Pr(>|z|)> 0.05 indicating that they are not significant in predicting the probability.

## Analyze the accuracy
```{r}
predict_log1<- predict(log_model1, type="response", validating_data)

validating_data$predicted1<- as.numeric(predict_log1>= 0.5)
table(validating_data$Survived, validating_data$predicted1)
model_accuray<- (119+53)/(119+53+32+18)*100
#The accuracy of the model is 77.48%

```



## Apply the model to test data 
```{r}
predict_test <- predict(log_model1, type="response", newdata= test_data)

test_data$predcted_survived <- as.numeric(predict_test> 0.5)
Titanic_predictions = data.frame(test_data[c("PassengerId","predcted_survived")])
write.csv(file = "TitanicPred", x = Titanic_predictions)

```



